\documentclass[12pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{fancyhdr}
\usepackage{parskip}
\usepackage[english]{babel}
\usepackage{amssymb}

\pagestyle{fancy}
\fancyhf{}
\rhead{Arnau Abella \- MIRI}
\lhead{Computational Complexity}
\rfoot{Page \thepage}

\title{Homework 1}
\author{Arnau Abella}
\date{\today}

\begin{document}

\section{Exercise: Composition of logspace computable functions}

A function $f : \{ 0,1 \} ^* \to \{ 0,1 \} ^*$ is called \textit{moderate} if there is exist a polynomial $p(n)$ such that $|f(x)| \leq p(|x|)$ for every $x \in \{ 0, 1 \} ^*$.


The \textit{bit-graph} of $f$ is the following language:


\[
BIT_f := \{  \langle x, i, b\rangle : 1 \leq i \leq |f(x)| \text{and the i-th bit of}\ f(x) \text{ is } b \}
\]


We say that $f$ is \textit{logspace computable} if $BIT_f$ is decidable in space $O(\log n)$; i.e., it is in L. Show that if $f : \{ 0,1 \} ^* \to \{ 0,1 \} ^*$ and $g : \{ 0,1 \} ^* \to \{ 0,1 \} ^*$ are both moderate and logspace computable, then the composition $f \circ g$ is also moderate and logspace computable.


\begin{proof}
  First, we are going to prove that \textit{moderate} is close under composition, and then, that \textit{logspace computability} is also close under composition.


  To prove that \textbf{if $f$ and $g$ are moderate, then $f \circ g$ is moderate} we are going to show a polynomial $p(n)\ s.t\ |f(g(x))| \leq p(|x|) \ \text{for every} x \in \{ 0, 1 \} ^*$.


  We apply $g(x)$ to $f$ and because $f$ is moderate, then there is a polynomial $p''(n)$ such that $|f(g(x))| \leq p''(|g(x)|)$. But, we also know that $g$ is moderate so for every $x \in \{ 0, 1 \} ^*$ there is a polynomial $p'(n)$ such that $|g(x)| \leq p'(|x|)$. Hence, $|f(g(x))| \leq p''(p'(|x|))$. If we reduce the polynomials to their \textit{canonical form} i.e.\ $p'(n) \equiv n^c $ and $p''(m) \equiv m^{c'} $, then $|f(g(x))| \leq (|x|^c)^{c'}$. Finally we replace $(|x|^c)^{c'} by n^d$ and replace it by its polynomial form $p(n)$ which result in $|f(g(x))| \leq p(|x|)$.


  To prove that \textbf{if $f$ and $g$ are logspace computable, then $f \circ g$ is logspace computable} we are going to show that $BIT_{f \circ g}$ is decidable in space $O(\log n)$. $BIT_{f \circ g}$ is decidable in space $O(\log n)$ means that there is a TM $M$ that computes the function $f_L: \{ 0,1 \} ^* \to  \{ 0,1 \} ^*$, where $f_L(x) = 1$ and at most $c \cdot \log n$ locations ($c$ is a constant) on M's work tapes are ever visited by $M$'s head during its computation on every input of length $n$. Let's define a TM $M$ that on input $\langle x, i, b \rangle$ checks if $f(g(x)) = b$ in the $i$-th bit in \textit{logspace} by using both $M_f$ and $M_g$ that are given by the statement of this problem.


  The algorithm is simple. Using $M_g$ we can compute $g(x)$, store the result in M's working tape, and then compute $f(g(x))$ using $M_f$. But this computation is not \textit{logspace computabe} even though the worktapes of $f$ and $g$ only uses $O(\log n)$ space. The problem is that we need to copy the output of $g(x)$ into one of M's worktapes in order to input it to $M_f$, and this output is \textbf{not} logspace.


  The trick is to notice is that, altough $M_f$ needs every bit of $g(x)$, it needs only \textbf{one bit at a time}. So we can compute $g(x)$ and copy a single bit to the work tape at each iteration of $M_f$ and discard the rest.


  To store the current iteration of $M_f$ we need a counter. The space of the counter can be limited to $O(\log n)$ because $1 \leq i \leq |f(g(x))|$ by the definition of $BIT_f$ and we proved that $|f(g(x))| \leq p(n)$ on the first part, so the \#iteration can be encoded in $c \cdot \log n$, where $c$ is a constant. Additionally, we need to know the length of $g(x)$ in order to stop processing $M_f$ and halt, however, this is not given, but it can be decided by computing both $M_g(\langle x, i, 0\rangle)$ and $M_g(\langle x, i, 1\rangle)$. If the output of both is 0, then $g(x)$ has finished and we can halt $M$.


  Our TM $M$ needs 4 tapes (input, output, counter, result of the $it$-th bit $g(x)$). We start by setting the counter to 0, computing $g(x)$ using $M_g(x, i, b)$ (we need to perform the test of halting we mentioned in the previous paragraph), copying the $n$th-bit $u$ ($n$ is given by the counter), and computing $M_f(u, i, b)$ (using the same $b$ that halted $M_g$) on that bit and copying the result to the output tape. This needs to be done recursively. At each new iteration we will increase the counter by one. As soon as $M_g$ halts, we can halt $M$ and the answer will be written in the output tape.


  This TM $M$ can decide the language $BIT_{f \circ g}$ for every input $\langle x, i \rangle$ by applying the same procedure of discarding the non-neeed bits of $g(x)$.


\end{proof}

\section{Exercise: Logspace verifiers}

Recall that NP has been characterized as the class of languages A for which there exists a polynomial $p(n)$ and a language $B$ in P such that for every string $x$ we have:

\[
  x \in A \Leftrightarrow \exists y \in \{ 0 , 1 \} ^ * \, s.t. \, |y| \leq p(|x|) \, and \, \langle x, y \rangle \in B.
\]

Show that P can be replaced by L and we still get a characterization of NP\; i.e.\ the verifier can be restricted to running not only in polynomial time but even logarithmic space (since L $\subseteq$ P , one half of this statement is obvious; you are asked to prove the otehr half without proving P $\subseteq$ L which, fyi, is an open problem).

\begin{proof}

  In order to prove $\implies$ we are going to use a similar technique that the one used in Cook-Levin Theorem proof. To proof it, it shows a polynomial-time Karp reducible function from any $L$ in NP and every $x \in L$ to $SAT$. The naive approach results into a $\psi_x$ CNF formulae of length $\ell 2^{\ell}$ which is \textbf{not} computable in polynomial time. The technique used in the proof to find such $\varphi_x$ that can be computed in polynomial time for every $x \in L$ is the following: if $L$ is in NP, then there exists a two-tape oblivious polynomial TM such that:

\begin{equation} \label{eq:1}
x \in L \Leftrightarrow \exists u \in \{ 0, 1 \}^{p(|x|)} \ s.t. \ M(x,u) = 1
\end{equation}

Then, it uses the property that the position of the head on an oblivious TM does not depend on the input, only of its size and the fact that this TM has only two tapes to translate this TM to a sequence of strings that uniquely determine this TM and it can be verified so by checking the following conditions:

\begin{enumerate}
  \item The first $n$ bits of $y$ are equal to $x$.
  \item The string $z_1$ encodes the initial snapshot of $M$. That is, $z1$, encodes the triple $\langle \triangleright, \square, q_{start} \rangle$.
  \item For every $i \in \{ 2, \dots , T(n) \}$, $z_i = F(z_{i-1}, z_{inputpos(i)}, z_{prev(i)})$.
  \item The last string $z_{t(n)}$ encodes a snapshot in which the machine halts and ouputs 1.
\end{enumerate}

  So, the formula $\varphi_x$ will take variables $y \in \{0,1\}^{n+p(n)}$ (the concatenation of the input $x$ and the certificate $u$)  and $z \in \{ 0, 1 \}^{cT(n)}$ (the snapshots/configuraitons) and will verify that y,z satisfy the $AND$ of these four conditions. Thus $x \in L \Leftrightarrow  \varphi_x \in SAT$. The last step of the proof consist of proving that we can express the four conditions of $\varphi_x$ in polynomial time.


  From the previous proof, we should realize that any oblivious TM can be translated into a finite series of snapshots and given $y$, $z$ (as defined above) we can verify that those snapshots uniquely encode the oblivious TM $M$ and they produce the same output as if you were running $M(x,u)$. We only need to proof that there is a TM that can verify the previous four conditions, not only in polynomial time as stated in Cook-Levin Theorem, but also in \textit{logspace}.


  Both 1,2,4 are easy to proof because any $z_i$ has at most length $c$ where c is a constant that depends on the characteristics of the TM. The tricky part is proving that 3 can be checked in logspace. The variable $i$ ranges from 1 to $T$, which is the total number of steps the $TM$ being snapshot takes before halting. Since the TM runs in polynomial time, i.e. $n^c$ for some constant $c$. This the number of bits needed to represent $i$ should only range from $1$ to $c \cdot \log n$ which is only logarithmic space. I should also add, that if you think about it, all that you need to compute the next snapshot is, the previous snapshot, $z_{i-1}$ which doesn't vary based on input size. Also the current input tape content which needs at most $\log n$ bits as an index, and having to remember when the head had last visited its current working tape position or index it's currently on to re-check its tape contents to determine the next move, which is at most a constant. Thus it can be done under logarithmic space.


  The $\Longleftarrow$ is trivial to proof.


\end{proof}

\end{document}
